\documentclass[11pt]{amsart}

\usepackage{a4wide}
\usepackage{paralist}
\usepackage{url}

\newcommand{\cA}{\mathcal{A}}
\newcommand{\cS}{\mathcal{S}}

\begin{document}
\begin{center}
\textbf{\sffamily
   Discrete and Algorithmic Geometry }

\medskip
   Julian Pfeifle,
   UPC, 2016 
\end{center}

\bigskip

\begin{center}
  \textbf{\sffamily Participants}
\end{center}

\medskip

\section*{Tim Weenink}
(TimWeenink) 

Hola! My name is Tim and I come from the Netherlands. I did my bachelor in mathematics at the TU Eindhoven and right now I'm spending the semester in Barcelona as an Erasmus student. The part of mathematics that I find most interesting is cryptology. Regarding this course, I hope we get to see some practical applications of the theory we learned in the first part. 

\medskip

\section*{Simon Van den Eynde}
(Simonvde)

Hello! I'm Simon and I'm from Belgium. This is my first year of the master and I'm also doing a semester here as an Erasmus student. At my university I followed some programming courses in which I learned to code Java decently and got an introductory course on C/C++. I also used python-based math packages (mainly Sagemath.org). For this course I don't have a specific thing in mind, but I liked what we did in the first semester and coding a thing or two on polygons would be interesting, for example using CGAL.

\medskip

\section*{Joaquim Casals Bu\~{n}uel}
Github login : \emph{quimFIB}\\

I'm a computer scientist graduate from the FIB. I spent last semester in the MIRI master but decided to switch master's after taking the
MAMME course ``Combinatorics''. I am mainly interested in Algorithms and Logic and in particular I really enjoy Complexity Theory. I took
this course because I had taken GEOC while an undergrad (the computational geometry course taught by Vera at the FIB) and I enjoyed it, so for this second
part hopefully I get to learn a thing or two about geometry in general. As for the future, I'm contemplating options, one of them being pursuing
a PhD in Complexity Theory / Logics.

\medskip

\section*{Next name here}
(another github login)

More of the same


\end{document}

